---
title: "Challenge-7"
author: "Loy Yee Keen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data: Palmer Penguins

Measurements for penguin species, island in Palmer Archipelago, size (flipper length, body mass, bill dimensions), and gender.

```{r, eval=TRUE, echo=FALSE}
library(tidyverse)
library(palmerpenguins)
glimpse(penguins)
```

# Palmer Penguins: Plot recreation

a. Start with the penguins data frame

```{r, eval=TRUE, echo=FALSE}
ggplot(data = penguins)
```

b. Map bill depth to the x-axis

```{r, eval=TRUE, echo=FALSE}
ggplot(data = penguins,
mapping = aes(x = bill_depth_mm))
```

c. Map bill depth to the y-axis

```{r, eval=TRUE, echo=FALSE}
ggplot(data = penguins,
mapping = aes(x = bill_depth_mm,
y = bill_length_mm))
```

d. Represent each observation with a point

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
mapping = aes(x = bill_depth_mm,
y = bill_length_mm)) +
geom_point()
```

e. Map species to the colour of each point

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
mapping = aes(x = bill_depth_mm,
y = bill_length_mm,
colour = species)) +
geom_point()
```

f. Title the plot "Bill depth and length"

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
mapping = aes(x = bill_depth_mm,
y = bill_length_mm,
colour = species)) +
geom_point() +
labs(title = "Bill depth and length")
```

g. Add the subtitle "Dimensions for Adelie, Chinstrap, and Gentoo Penguins"

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
mapping = aes(x = bill_depth_mm,
y = bill_length_mm,
colour = species)) +
geom_point() +
labs(title = "Bill depth and length",
subtitle = "Dimensions for Adelie,
Chinstrap, and Gentoo Penguins")
```

h. Label the x and y axes as "Bill depth (mm)" and "Bill length (mm)", respectively

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
 mapping = aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species)) +
 geom_point() +
 labs(title = "Bill depth and length",
subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
 x = "Bill depth (mm)",
 y = "Bill length (mm)")
```

i. Label the legend "Species"

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
 mapping = aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species)) +
 geom_point() +
 labs(title = "Bill depth and length",
 subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
 x = "Bill depth (mm)", y = "Bill length (mm)",
 colour = "Species")
```

j. Add a caption for the data source

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
 mapping = aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species)) +
 geom_point() +
 labs(title = "Bill depth and length",
 subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
 x = "Bill depth (mm)", y = "Bill length (mm)",
 colour = "Species",
 caption = "Source: Palmer Station LTER/palmerpenguin.package")
```

k. Finally, use a discrete colour scale that is designed to be perceived by viewers with common forms of colour blindness.

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
 mapping = aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species)) +
 geom_point() +
 labs(title = "Bill depth and length",
 subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
 x = "Bill depth (mm)", y = "Bill length (mm)",
 colour = "Species",
 caption = "Source: Palmer Station LTER/palmerpenguin.package") +
 scale_colour_viridis_d()
```
Can omit the names of first two arguments when building plots with ggplot()

```{r, eval=TRUE, echo=FALSE} 
ggplot(data = penguins,
 mapping = aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species)) +
 geom_point() +
 scale_colour_viridis_d()
```

vs

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins) + # Data layer
 aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species) + # Aesthetics layer
 geom_point() + # Geometric layer
 scale_colour_viridis_d()
```

vs

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins,
 aes(x = bill_depth_mm,
 y = bill_length_mm,
 colour = species)) +
 geom_point() +
 scale_colour_viridis_d()
```

# Palmer Penguins: Colour

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins) + aes(x = bill_depth_mm, y = bill_length_mm,
 colour = species) +
 geom_point() + scale_colour_viridis_d()
```

# Palmer Penguins: Shape

Island

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species,
 shape = island)) +
 geom_point() + scale_colour_viridis_d()
```

Species

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species,
 shape = species)) +
 geom_point() + scale_colour_viridis_d()
```

Size

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species, shape = species,
 size = body_mass_g)) +
 geom_point() + scale_colour_viridis_d()
```

# Alpha

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species,
 shape = species, size = body_mass_g, alpha = flipper_length_mm)) +
 geom_point() + scale_colour_viridis_d()
```

# Mapping vs Setting

Mapping

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins) +
 aes(x = bill_depth_mm,
 y = bill_length_mm,
 size = body_mass_g,
 alpha = flipper_length_mm) +
 geom_point()
```

Setting

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins) +
 aes(x = bill_depth_mm,
 y = bill_length_mm) +
 geom_point(size = 2, alpha = 0.5)
```

# Faceting

```{r, eval=TRUE, echo=FALSE} 
ggplot(penguins) +
 aes(x = bill_depth_mm,
 y = bill_length_mm) +
 geom_point() +
 facet_grid(species ~ island)
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
 facet_grid(species ~ sex)
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
 facet_grid(sex ~ species)
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
 facet_wrap(~ species)
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
 facet_wrap(~ species, ncol = 2)
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
 facet_grid(. ~ species)
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
 geom_point() + facet_grid(species ~ sex) + scale_color_viridis_d()
```

```{r, eval=TRUE, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
 geom_point() + facet_grid(species ~ sex) + scale_color_viridis_d() +
 guides(color = "none")
```

# Data 2: Lending Club

Take a peek at data

```{r, eval=TRUE, echo=FALSE}
library(openintro)
glimpse(loans_full_schema)
```

Selected variables

```{r, eval=TRUE, echo=FALSE}
loans <- loans_full_schema %>%
 select(loan_amount, interest_rate, term, grade,
 state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```

# Histogram

```{r, eval=TRUE, echo=FALSE}
ggplot(loans) + aes(x = loan_amount) +
 geom_histogram()
```

Set Binwidth to 1000

```{r, eval=TRUE, echo=FALSE}
# binwidth = 1000
ggplot(loans, aes(x = loan_amount)) +
 geom_histogram(binwidth = 1000)
```

Set Binwidth to 5000

```{r, eval=TRUE, echo=FALSE}
# binwidth = 5000
ggplot(loans, aes(x = loan_amount)) +
 geom_histogram(binwidth = 5000)
```

Set Binwidth to 20000

```{r, eval=TRUE, echo=FALSE}
# binwidth = 20000
ggplot(loans, aes(x = loan_amount)) +
 geom_histogram(binwidth = 20000)
```

# Customising Histograms

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount)) + geom_histogram(binwidth = 5000) +
 labs(x = "Loan amount ($)", y = "Frequency", title = "Amounts of Lending Club loans") 
```

# Fill with a categorical variable

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) +
 geom_histogram(binwidth = 5000, alpha = 0.5) +
 labs(x = "Loan amount ($)",y = "Frequency",title = "Amounts of Lending Club loans")
```

# Facet with a categorical variable

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) + geom_histogram(binwidth = 5000) +
 labs(x = "Loan amount ($)",y = "Frequency",title = "Amounts of Lending Club loans") +
 facet_wrap(~ homeownership, nrow = 3)
```

# Density plot

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount)) +
 geom_density()
```

# Density plots and adjusting bandwidth

Adjust to 0.5

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount)) +
 geom_density(adjust = 0.5)
```

Adjust to 1

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount)) +
 geom_density(adjust = 1) # default bandwidth
```

Adjust to 2

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount)) +
 geom_density(adjust = 2)
```

# Customising density plots
 
```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount)) +
 geom_density(adjust = 2) +
 labs( x = "Loan amount ($)", y = "Density", title = "Amounts of Lending Club loans" )
```

# Adding a categorical variable

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) +
 geom_density(adjust = 2, alpha = 0.5) +
 labs(x = "Loan amount ($)",y = "Density",title = "Amounts of Lending Club loans", fill = "Homeownership")
```

#Box Plot

Box Plot of Interest Rate

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = interest_rate)) +
 geom_boxplot()
```

Box Plot of Annual Income

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = annual_income)) +
 geom_boxplot()
```

#Customising Box Plots

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = interest_rate)) +geom_boxplot() +labs(x = "Interest rate (%)",y = NULL,
 title = "Interest rates of Lending Club loans") +
 theme( axis.ticks.y = element_blank(), axis.text.y = element_blank() )
```

#Adding a categoric variable

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = interest_rate,
 y = grade)) +
 geom_boxplot() +
 labs(x = "Interest rate (%)",y = "Grade",title = "Interest rates of Lending Club loans",subtitle="by grade of loan")
```

# Scatterplot

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = debt_to_income, y = interest_rate)) +
 geom_point()
```

#Hex Plot

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = debt_to_income, y = interest_rate)) +
 geom_hex()
```

```{r, eval=TRUE, echo=FALSE}
ggplot(loans %>% filter(debt_to_income < 100),
 aes(x = debt_to_income, y = interest_rate)) +
 geom_hex()
```
 
# Bar Plot

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = homeownership)) +
 geom_bar()
```

Segmented Bar Plot

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = homeownership,
 fill = grade)) +
 geom_bar()
```

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = homeownership, fill = grade)) +
 geom_bar(position = "fill")
```

# Customising Bar Plots

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(y = homeownership, fill = grade)) + geom_bar(position = "fill") +
labs( x = "Proportion", y = "Homeownership", fill = "Grade", title = "Grades of Lending Club loans", subtitle="and home ownership of lendee")
```

# Violin Plots

```{r, eval=TRUE, echo=FALSE}
ggplot(loans, aes(x = homeownership, y = loan_amount)) +
 geom_violin()
```

# Ridge Plots

```{r, eval=TRUE, echo=FALSE}
library(ggridges)
ggplot(loans, aes(x = loan_amount, y = grade, fill = grade, color = grade)) +
 geom_density_ridges(alpha = 0.5)
```
