---
title: "Challenge-6"
author: "Loy Yee Keen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions
<br>

#### Question-1: Countdown Blastoff (While Loop)

Create a program that simulates a rocket launch countdown using a while loop. Start from 10 and countdown to "Blastoff!" with a one-second delay between each countdown number. Print a message when the rocket launches.

**Hint:** You may want to use `cat` command to print the countdown and `Sys.sleep` for incorporating the delay

**Output preview:** Here is how the countdown could look like 

**Solutions:**

```{r, eval=T,echo=TRUE}
# Enter code here

i <- 10

while (i >= 0) {
  if (i == 10) {
    cat("10\n")
  } else if (i == 0) {
    cat("Blastoff!\n")
  } else {
    cat(i, "\n")
  }
  Sys.sleep(1)  
  i <- i - 1
}

      
```
<br>

#### Question-2: Word Reverser (for Loop)

Develop a program that takes a user-entered word and uses a while loop to print the word's characters in reverse order. For example, if the user enters "hello," the program should print "olleh."

**Hint:** You may want to use `substr` command to access each character of the input word, and `paste` command to join the reversed letters one at a time

**Solutions:** 

```{r, eval=T,echo=TRUE}
# Enter code here

word <- readline("Enter a word: ")

i <- nchar(word)
reversed_word <- character(1)
while (i > 0) {
  char <- substr(word, i, i)
  reversed_word <- paste0(reversed_word, char, sep = "")
  i <- i - 1
}

cat("Reversed word:", reversed_word, "\n")

```
<br>
